Visual narratives of the Covid-19 pandemic

### 2022-01-12 JDSSV_manuscript_knitr.Rnw

%% -- Introduction -------------------------------------------------------------

@@

@@

>>>>

<<get-nyt-data, include = F>>=

# Read in the data from NYT
library(tidyverse)
library(lubridate)
library(zoo)
library(ggrepel)
library(gridExtra)
cvstate <- read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv")

cvstate <- cvstate %>%
  group_by(state, fips) %>%
  mutate(date_100th_case = min(date[cases >= 100]),
         days_100th_case = date - date_100th_case,
         daily_cases = cases - lag(cases, 1, default = 0),
         week_avg_cases = rollmean(cases, 7, align = "right", fill = c(0, NA, NA), na.pad = T, na.rm = T),
         daily_avg_cases = rollmean(daily_cases, 7, align = "right", fill = c(0, NA, NA), na.pad = T, na.rm = T),
         daily_deaths = deaths - lag(deaths, 1, default = 0),
         week_avg_deaths = rollmean(deaths, 7, align = "right", fill = c(0, NA, NA), na.pad = T, na.rm = T),
         daily_avg_deaths = rollmean(daily_deaths, 7, align = "right", fill = c(0, NA, NA), na.pad = T, na.rm = T)) %>%
  ungroup() %>%
  mutate(state = factor(state))

download.file("https://www2.census.gov/programs-surveys/popest/tables/2010-2019/state/totals/nst-est2019-01.xlsx", destfile = "census_pop_state.xlsx", quiet = T, mode = "wb")
pop_data <- readxl::read_xlsx("census_pop_state.xlsx", col_names = c("state", "pop2019"), col_types = c("text", rep("skip", 11), "numeric"), skip = 9) %>%
  mutate(state = str_remove(state, "[[:punct:]]"))

cvstate <- left_join(cvstate, pop_data, by = "state")

my_theme <- theme_bw() + 
  theme(plot.background = element_rect(fill = "transparent", color = NA), 
        panel.background = element_rect(fill = "transparent", color = NA), 
        legend.background = element_rect(fill = "transparent", color = NA),
        axis.text.y = element_text(angle = 90, hjust = 0.5)) 
theme_set(my_theme)
@

<<colorpal, include = F>>=
# Including this for temporary compatibility - we may want to use *shudder* viridis?
main_colors <- c("#d00000", "#f5f1e7", "#c7c8ca")

# use as accents
secondary_colors <- c("#001226", "#249ab5")

# These are only to be used for infographics/charts
tertiary_colors <- c("#bccb2a", "#f58a1f", "#005d84", "#ffd74f", "#a5228d")

color_palette <- c("#d00000", "#249ab5",  "#1B8500", "#f58a1f", "#a5228d", "#001226", "#B1EB66", "#005d84", "#c25700")

heat_color_palette <- c("#ffd74f", "#f58a1f", "#c25700", "#d00000", "#9D0000", "#6A0000")

blueseqpal <- c("001226", "06293E", "0C3F56", "12566E", "186D85", "1E839D", 
                "249AB5", 
                "39ABC1", "4EBCCE", "64CDDA", "79DDE6", "8EEEF3", "A3FFFF") %>% 
  paste0("#", .) %>% rev()

blueseqpal_lt <- rev(c("#B2FFFF", "#C2FFFF", "#D1FFFF", "#E0FFFF", "#F0FFFF"))

darkblueseqpal <- colorspace::darken(blueseqpal, .2)
lightblueseqpal <- colorspace::lighten(blueseqpal, .2)
@
%% -- Introduction -------------------------------------------------------------

@@

<<freq-domain-plot, include = F, eval = F>>
show_states <- c("California", "Texas", "New York", "Arizona", "Iowa", "Washington")

n <- 50000
tmp <- cvstate %>%
  filter(state %in% show_states) %>%
  select(state, date, cases) %>%
  mutate(mod_num = cases %% n,
         whole_num = floor(cases/n)) %>%
  group_by(state, whole_num) %>%
  mutate(whole_num_days = n()) %>%
  group_by(state) %>%
  arrange(date)

tmp2 <- bind_rows(
  # Day before first case detected:
  filter(tmp, whole_num == 0) %>%
    group_by(state) %>%
    summarize(date = min(date) - lubridate::days(1), y = 0, whole_num = 0, 
              whole_num_days = unique(whole_num_days), mod_num = 0) %>%
    unique() %>%
    ungroup(),
  # Date before the count rolls over
  tmp %>%
    group_by(state, whole_num) %>%
    filter(date == max(date)) %>%
    group_by(state) %>%
    mutate(y = ifelse(whole_num == max(whole_num), mod_num, n),
           whole_num_days = ifelse(whole_num == max(whole_num), whole_num_days/(y/10000), whole_num_days)),
  # Date after the count rolls over
  tmp %>%
    group_by(state, whole_num) %>%
    filter(date == min(date)) %>%
    mutate(y = 0, date = date - lubridate::days(1))
  ) %>%
  group_by(state, whole_num) %>%
  mutate(whole_num_days = max(whole_num_days))

ggplot(data = tmp2, aes(x = date, ymin = 0, ymax = y, 
                        color = whole_num_days, 
                        fill = whole_num_days, 
                        group = whole_num)) + 
  geom_ribbon() + 
  facet_grid(state~.) + 
  scale_x_date(date_breaks = "month", date_labels = "%B") +
  scale_y_continuous(breaks = c(n/2, n), labels = paste0(c(n/2000, n/1000), "k"))  + 
  scale_color_gradientn(guide = 'none', paste0("Days to ", n/1000, "k\nmore cases"), space = "Lab", 
    trans = "log10",
    colours = darkblueseqpal
    ) + 
  scale_fill_gradientn(paste0("Days to ", n/1000, "k\nmore cases"), space = "Lab", 
    trans = "log10",
    colours = blueseqpal) + 
  my_theme + 
  theme(axis.title = element_blank(), 
        legend.background = element_rect(fill = "transparent", color = NA), 
        legend.position = c(0, .66), legend.justification = c(0, 1), legend.direction = "horizontal")
@

@@

<<log-scale-failures>>=
# ----- Set up labeller to tweak things ----------------------------------------
rlistmap <- function(x, myfun) {
  if ("list" %in% class(x)) return(purrr::map(x, ~rlistmap(., myfun)))

  return(myfun(x))
}
fix_labels <- function(x) {
  x  %>%
    str_replace('scale: (.*)', "\\1 scale") %>%
    str_replace("baseline: (\\d{1,})", "Baseline:\n\\1 cases")
}
label_custom <- function(labels, multi_line = T, sep = ": ") {
  out <- label_both(labels, multi_line = multi_line, sep = sep)

  out <- rlistmap(out, fix_labels)

  out
}
class(label_custom) <- "labeller"

# ----- Set up data ------------------------------------------------------------
growth <- rpois(30, 15*exp((1:30 - 20)/5))

tmp <- tibble(baseline = 15,
       x = 1:30,
       y = rpois(30, 15),
       y1 = growth) %>%
  bind_rows(
    tibble(baseline = 50,
       x = 1:30,
       y = rpois(30, 50),
       y1 = growth) 
  ) %>%
  bind_rows(
    tibble(baseline = 200,
       x = 1:30,
       y = rpois(30, 200),
       y1 = growth) 
  ) %>%
  pivot_longer(matches("y")) %>%
  mutate(name = str_replace_all(name, c("y$" = "Background", "y1" = "Exponential")))  %>%
  group_by(baseline, x) %>%
  mutate(Total = sum(value)) %>%
  pivot_wider(names_from = "name", values_from = "value") %>%
  pivot_longer(matches("Background|Exponential|Total")) %>%
  mutate(name = factor(name, levels = c("Total", "Exponential", "Background")))

tmp <- bind_rows(mutate(tmp, scale = "Log"), mutate(tmp, scale = "Linear")) %>%
  mutate(Baseline = reorder(paste0("Baseline: ", baseline, " cases"), baseline)) %>%
  mutate(scale = factor(scale, levels = c("Log", "Linear")))

# ----- Set up scales ----------------------------------------------------------
library(facetscales)
scales_y <- list(
  Log = scale_y_log10(),
  Linear = scale_y_continuous()
)

tmp %>%
ggplot(aes(x = x, y = value, color = name, size = name)) + 
  geom_line(alpha = .9) +
  facet_grid_sc(rows = vars(scale), cols = vars(Baseline), 
                scales = list(x = "fixed", y = scales_y)) +
  scale_size_manual(values = c(1.75, 1, 1)) +
  scale_color_manual(values = c("#249ab5", "#d00000", "#001226")) +
  scale_fill_discrete(name = "") + 
  theme(legend.position = "bottom", axis.title = element_blank(),
        legend.title = element_blank())
@
